### 2022-01-19 proposal.Rmd
added proposal

---
title: "Project Proposal"
author: "Peyton Kuhlers"
date: "1/18/2022"
output: pdf_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Patient Derived Xenografts (PDX) are *in vivo* cancer models
in which patient-derived tumors are taken and engrafted into immuno-deficient mice.
Once established, PDXs can be used as platforms for drug discovery or
investigation of *in vivo* drug mechanisms.

The PDX Development and Trial Centers Research Network (PDXNet) is an
NCI consortium that has generated publicly available for
PDX models across a large number of cancer types.
This data includes genomic characterizations and tumor growth measurements for
a variety of drugs.

My goal is to analyze the drug response data of one or two
cancers to determine if any of the drugs have a significant effect 
on the growth of tumors in PDX models.
Growth data for colon cancer includes ~2700 volume measurements,
7 drugs (and 1 control), and 263 mice originating from 7 unique patients.

Because there is also a wealth of genomic data available as well,
I may incorporate data such as tumor mutation or subtype --if feasible--
to identify any mutation or subtype specific drug responses.