### 2022-02-02 proposal.Rmd
updated proposal

---
title: "Project Proposal"
author: "Peyton Kuhlers"
output: pdf_document
geometry: "left=2cm,right=2cm,top=0.25cm,bottom=1.5cm"
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Patient Derived Xenografts (PDX) are *in vivo* cancer models
in which patient-derived tumors are taken and engrafted into immuno-deficient mice.
Once established, PDXs can be used as platforms for drug discovery or
investigation of *in vivo* drug mechanisms (1).
The PDX Development and Trial Centers Research Network (PDXNet) is an
NCI consortium that has generated publicly available for
PDX models across a large number of cancer types.
This data includes genomic characterizations and tumor growth measurements for
a variety of drugs.

Evrard et al (2) used multiple strategies to evaluate tumor response
from three PDX models treated with temozolomide.
These included percent change in tumor volume,
area under the curve, and progression free survival, among others,
but did not use any linear models of the volume measurements directly.
Pearson et al (3) characterized the growth rates of PDX tumors
across passages using linear mixed models and noted that
tumor volumes on the log-scale seem to grow quadratically.

My proposal is to analyze the drug response data
available from the PDXnet for one or two
cancers to determine if any of the tested drugs significantly inhibit
tumor growth.
Data will be downloaded from the PDXNet online portal
(https://portal.pdxnetwork.org/).
Growth data for colon cancer includes ~2700 volume measurements,
7 drugs (and 1 control), and 263 mice originating from 7 unique patients.
I plan to use a linear model fit with generalized least squares
to determine if any treatments slow tumor growth.
Additionally, I plan to use a mixed model approach to
investigate the variation in response to treatment.

The PDXNet portal also contains genomic data,
such as tumor mutational burden (TMB)
and homologous recombination defficiency analysis (HRD),
which may impact response to treatment (4,5).
I aim to incorporate TMB and HRD measurements into the linear models and
test for their effect on tumor growth and interaction with treatment.

References
----------
1) Tentler, J., Tan, A., Weekes, C. et al. Patient-derived tumour xenografts as models for oncology drug development. Nat Rev Clin Oncol 9, 338–350 (2012). https://doi.org/10.1038/nrclinonc.2012.61
2) Evrard, Yvonne A et al. “Systematic Establishment of Robustness and Standards in Patient-Derived Xenograft Experiments and Analysis.” Cancer research vol. 80,11 (2020): 2286-2297. doi:10.1158/0008-5472.CAN-19-3101
3) Pearson, Alexander T., et al. "Patient-derived xenograft (PDX) tumors increase growth rate with time." Oncotarget [Online], 7.7 (2016): 7993-8005. Web. 2 Feb. 2022 
4) Yarchoan, Mark et al. “Tumor Mutational Burden and Response Rate to PD-1 Inhibition.” The New England journal of medicine vol. 377,25 (2017): 2500-2501. doi:10.1056/NEJMc1713444
5) Aguirre, Andrew J et al. “Real-time Genomic Characterization of Advanced Pancreatic Cancer to Enable Precision Medicine.” Cancer discovery vol. 8,9 (2018): 1096-1111. doi:10.1158/2159-8290.CD-18-0275