### 2022-03-02 proposal_presentation.Rmd
presentation update

---
title: "Analysis of PDX Growth Data"
author: "Peyton Kuhlers"
date: "3/2/2022"
output: 
  beamer_presentation:
    theme: "AnnArbor"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Background

- Patient Derived Xenografts (PDX) are *in vivo* cancer models used as platforms
for drug discovery, biomarker identification, and mechanistic studies\footnote{Tentler et al 2012}

- Becoming preferred model as they preserve donor genomic characteristics
and allow for microenvironment modelling *in vivo*

\centering
\includegraphics[width = 8cm, height = 4cm]{figures/pdx_graphic.png}

\note{In contrast to cells in a dish}

## Background Cont.

- The PDXNet is an NCI-sponsored consortium aimed at using PDX models
for high-throughput development and testing of cancer therapies

- Much work has centered around genomic characterization of PDX models\footnote{Sun et al 2021}\textsuperscript{,}\footnote{Woo et al 2020}

- While data is available, comparatively less work has focused on 
analysis of tumor growth data in PDX models.

\note{Both in PDXnet and in the literature in general}

## Previous Work

- PDXNet collaborators\footnote{Evrard et al 2020} tested multiple
different strategies for analyzing tumor growth under treatment
  + Percent change
  + Modified RECIST
  + Area Under the Growth Curve
  + "Tumor growth inhibition"

\note{Lots of ignored data, not modelling the entire time course}
\note{Plus multiple clusters, mouse, patient, etc}

- In contrast, Pearson\footnote{Pearson et al 2016} and Oberg\footnote{Oberg et al 2021}
used mixed models to analyze growth over multiple passages and treatment effects, respectively.

## Significance

- Given PDXs are meant for high-throughput discovery,
robust and **efficient** methods are necessary
  + Inefficient statistical methods waste resources

## Data

- Growth data for colon cancer includes ~2700 volume measurements,
7 drugs (and 1 control), and 263 mice originating from 7 unique patients.

\centering
\includegraphics[width = 9cm, height = 5.5cm]{figures/growth_plots.pdf}

## Analysis

- Mixed effects model of tumor growth

\centering
$E[Y_{i}|b_i] = \beta_0 + b_{i1} + \beta_1(time_{i}) + b_{i2}(time) + \gamma_i (agent_i*time)$

## Result Reporting

- Profile Plots (shown above)

- Forest plot of interaction coefficients

- Table of specific pairwise contrasts between different therapies

- Variance of random effects

- As a secondary analysis, test methods from Evrard et al 
at e.g. day 75 and compare to model contrasts at day 75.

##

Thank You! Questions?