### 2022-04-06 paper_Kuhlers.Rmd
starting paper

---
title: "Analysis of Patient Derived Xenograft Growth Curves"
author: "Peyton Kuhlers"
date: "4/6/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

Introduction
=============

Patient Derived Xenografts (PDX) are *in vivo* cancer models
in which patient tumors are removed and engrafted into immuno-deficient mice.
Once established, PDXs can be used as platforms for drug discovery or
investigation of *in vivo* drug mechanisms (**REFS**).

PDX models have significant advantages over alternative models.
Standard *in vitro* models typically undergo transformations
to ensure proper growth and are kept in growth media that is drastically
different from the tumor enviroment *in vivo*. Additionally, these models
are "two dimensional" and all three dimensional interactions are
lost with cell lines (**REFS**).
PDX models preserve the 3D nature of human tumors and do not require
transformation, thereby also preserving the genomic characteristics of the
donor tissue (**REFS**).

Methods
=======

Tumor volume data was obtained from the PDXnet Portal,
queried for all colon adenocarcinoma.

Results
=======

\begin{centering}
\includegraphics[width = 6in, height = 8in]{figures/growth_plots.pdf}
\end{centering}

Discussion
===========