### 2022-04-21 paper_Kuhlers.Rmd
more writing

---
title: "Model Based Analysis of Patient Derived Xenograft Growth Curves"
author: "Peyton Kuhlers"
date: "`r Sys.Date()`"
output:
  pdf_document:
    extra_dependencies: ["setspace"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

\doublespacing

Introduction
=============

Patient Derived Xenografts (PDX) are *in vivo* cancer models
in which patient tumors are removed and engrafted into immuno-deficient mice.
Once established, PDXs can be used as platforms for drug discovery or
investigation of *in vivo* drug mechanisms (**REFS**).

PDX models have significant advantages over alternative models.
Standard *in vitro* models typically undergo transformations
to ensure proper growth and are kept in growth media that is drastically
different from the tumor enviroment *in vivo*. Additionally, these models
are "two dimensional" and all three dimensional interactions are
lost with cell lines (**REFS**).
PDX models preserve the 3D nature of human tumors and do not require
transformation, thereby also preserving the genomic characteristics of the
donor tissue (**REFS**).

Methods
=======

### Data Acquisition and Processing

Tumor volume data was obtained from the PDXnet Portal,
queried for all available colon adenocarcinoma.
All data were contributed from the NCI Patient-Derived Models Repository (PDMR).
Mice were staged to a tumor volume of 200mm^3^ (Evrard et al, NCI SOP, Alley et al)
before being randomized to a treatment group.
Tumors were measured by caliper 1-3 times per week,
potentially more depending on the health of animal (Evrard et al, NCI SOP).
The exact study endpoints for this data were not documented,
however Evrard et al performed a similar study with data from the PDMR,
where the study endpoints were "Animal health, Max. tumor size, 4000mm^3^,
or 300 days, if max. TV not reached" (See Table 1 of Evrard et al).
Additionally, the PDMR standard operating procedures
recommend harvesting tumors at 1000-2000mg.
The distribution of maximum tumor volumes in this dataset reflects this.
Observation days were rescaled to the number of days the sample
was observed, so that the time origin was 0 for all samples.
For example, if the mouse was staged on observation day 30,
this became day 0, and all subsequent measurements on this mouse were scaled to
days since day 30. Days were then rescaled to months.

### Mixed Effects Model

Tumor volumes were analyzed using generalized linear mixed effects models.
Fixed effects included were time and the time-drug interaction.
The drug main effect was not included to model equal mean volume
at baseline amongst all treatment groups.
A random intercept and random time slope were included for each mouse.

### "Naive" Accelerated Failure Time Model

In addition to analyzing tumor volumes directly, time-to-event was
also analyzed. Since only 2 mice achieved 4000mm^3^,
the volume endpoint was readjusted 1500mm^3^,
which was the 25^th^ percentile of volumes in this study.
A censoring indicator was constructed for whether or not
1500mm^3^ was achieved.
Time was taken as the day with the closest measurement to 1500mm^3^.
With these data, a Weibull accerated failure time (AFT) model was fit.

### "Imputed" Accelerated Failure Time Model

Since some mice exceed 1500mm^3^ (potentially by a substantial amount)
a "model imputed" time-to-event analysis was also considered.
If, for example, a mouse achieved 1500mm^3^ on day 30,
but was not measured until day 40 at 1700mm^3^,
we can consider time as "measured with error" and will
over-estimate the treatment effect.
Conversely, if we only measure 1000mm^3^ on day 20,
we will be under-estimating the treatment effect.
To address this, the above mixed effects model was solved
for time, using the mouse-specific random effects
to obtain a "model-imputed" time to 1500mm^3^ for each mouse.
These model imputed estimates were then used to fit
a Weibull AFT model. For this analysis it was assumed all mice reached
1500mm^3^ (i.e. no censoring).

Results
=======

\begin{centering}
\includegraphics[width = 6in, height = 8in]{figures/growth_rescaled_days.pdf}
\end{centering}

\begin{centering}
\includegraphics[width = 6in, height = 8in]{figures/mixed_forest.pdf}
\end{centering}

\begin{centering}
\includegraphics[width = 6in, height = 8in]{figures/aft_models_overlaid_forest.pdf}
\end{centering}

Discussion
===========