### 2022-04-26 presentation_Kuhlers.Rmd
Final presentation

---
title: "Model Based Analysis of Patient Derived Xenograft Tumor Volumes"
author: "Peyton Kuhlers"
date: "4/25/2022"
output:
  beamer_presentation:
    theme: "AnnArbor"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Background

- Patient Derived Xenografts (PDX) are *in vivo* cancer models used as platforms
for drug discovery, biomarker identification, and mechanistic studies\footnote{Tentler et al 2012}

- Becoming preferred model as they preserve donor genomic characteristics
and allow for microenvironment modelling *in vivo*

\centering
\includegraphics[width = 8cm, height = 4cm]{figures/pdx_graphic.png}

\note{In contrast to cells in a dish}

## Background Cont.

- The PDXNet is an NCI-sponsored consortium aimed at using PDX models
for high-throughput development and testing of cancer therapies

- Much work has centered around genomic characterization of PDX models\footnote{Sun et al 2021}\textsuperscript{,}\footnote{Woo et al 2020}

- While data is available, comparatively less work has focused on 
analysis of tumor growth data in PDX models.

\note{Both in PDXnet and in the literature in general}

## Previous Work

- PDXNet collaborators\footnote{Evrard et al 2020} tested multiple
different strategies for analyzing tumor growth under treatment
  + Percent change
  + Modified RECIST
  + Area Under the Growth Curve
  + "Tumor growth inhibition"

- In contrast, Pearson\footnote{Pearson et al 2016} and Oberg\footnote{Oberg et al 2021}
used mixed models to analyze growth over multiple passages and treatment effects, respectively.

## Significance

- Given PDXs are meant for high-throughput discovery,
we want to use methods that use all available data and maximize the
chance at finding an effect, while minimizing resource use

## Data

- ~2600 volume measurements, 7 drugs (and 1 control),
and 263 mice originating from 7 unique colon cancer patients.

- Generated by the NCI Patient Derived Models Repository (PDMR)

- Mice 'staged' to 150-200mm^3^ then randomized to treatment

\centering
\includegraphics[width = 7.5cm, height = 5cm]{figures/growth_rescaled_days.pdf}

## Proposed Models

- Mixed effects model of growth over time

- "Naive" accelerated failure time model

- "Imputed" accelerated failure time model

## Mixed Effects Model

- Is the growth rate significantly different between treatment and control?

\begin{center}
$log(E[Y_i|b_i]) = \beta_0 + b_{i1} + \beta_1(time_{i}) + b_{i2}(time_i) + \gamma_i (agent_i*time_i)$
\end{center}

- $Y_i$ is the volume for the i^th^ mouse - assumed Normal
- Interested in inference on $\gamma_i$

## Mixed Effects Model - Results

- Drug-time interactions -- estimate the change in treatment vs control over time
- Gemcitabine and Irinotecan weakly reduce growth rates relative to control

\centering
\includegraphics[width = 8cm, height = 5.5cm]{figures/mixed_forest.pdf}

## Accelerated Failure Time Models

- Do treatments extend the time to some volume endpoint?
  + PDMR uses 4000mm^3^ as endpoint (really big) - in practice 1000-2000mm^3^

\begin{center}
$log(T_i) = \beta_0 + \beta_1(agent_i) + \sigma\epsilon$
\end{center}

- $T_i$ is the (possibly censored) time to 1500mm^3^ for the i^th^ mouse, assumed Weibull
- Interested in inference on $\beta_1$

## Accelerated Failure Time Models - Cont.

- Time is measured "with error"
  + "Naive" model - use the time with volume closest to 1500
  + "Imputed" model - impute time to 1500 by solving the mixed effects model for time

## Accelerated Failure Time Models - Results

- Acceleration factors - Ratio of the mean time to 1500 in treatment vs control
- Imputed time does not seem to dramatically change estimates

\centering
\includegraphics[width = 8cm, height = 5.5cm]{figures/aft_models_overlaid_forest.pdf}

## Conclusions

- Gemcitibine and irinotecan had weak effects on growth rates relative to control
- All drugs (except Vemurafenib) extended the mean time to 1500mm^3^ by ~ 1.3 times
  + Equal mean at baseline assumption?
- Imputing time with a mixed model does not dramatically change acceleration estimates

## Limitations

- Exact details of the experiments unknown
  + Gleaned information from NCI SOPs and other publications
  + Could be these strains are all resistant to treatment

- More complex relationships were likely missed using a log-linear model
  + Some drugs may have more complex (nonlinear) effects on growth

- Addressed the naive AFT model with an imputed AFT model
  + Standard errors of imputed AFT are likely incorrect - bootstrap
  + Informative censoring?

- The number of times a mouse is measured is related to its well-being
  + Could produce biased estimates

##

Thank You! Questions?